---
title: 'Notebook #3'
author: "Simon Garnier"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
    fig_height: 6
    fig_width: 9
    number_sections: yes
    theme: cosmo
    toc: yes
---

--- 

# Data

```{r, message=FALSE}
# load required libraries
if (!require(pacman)) install.packages("pacman")
p_load(readr, dplyr, broom, ggplot2, cowplot, sjPlot, lme4)

# load data
num2 <- read_csv("../../data/num2.csv")
num2
  
prop2 <- read_csv("../../data/prop2.csv")
prop2

num4 <- read_csv("../../data/num4.csv")
num4
  
prop4 <- read_csv("../../data/prop4.csv")
prop4
```

[Back to top]("#")

---

# Day 2

## Number of blades cleared

### Graph

```{r, warning=FALSE, message=FALSE, fig.height=6, fig.width=6}
ggplot(num2, aes(x = Length, y = Numbercleared, color = Hardness, group = Hardness)) +
  geom_jitter(alpha = 0.25, width = 0.05) +
  stat_summary(fun.y = mean, geom = "point", size = 4) +
  stat_summary(fun.y = mean, geom = "line") +
  ylab("Number of blades cleared") +
  scale_color_brewer("Hardness", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```

---

### Model selection 

Note: a first pass at the model showed that the residuals weren't very clean. 
I used a square root transformation to correct some of it. 

```{r, warning=FALSE, message=FALSE}
m_full_num_2 <- lmer(sqrt(Numbercleared) ~ Length * Hardness + (1 | ColonyID), data = num2, REML = FALSE)
m_no_interaction_num_2 <- lmer(sqrt(Numbercleared) ~ Length + Hardness + (1 | ColonyID), data = num2, REML = FALSE)
anova(m_full_num_2, m_no_interaction_num_2)
```

There is a statistical difference between the full model and the model without 
interaction.We'll continue with the full model. 

---

### Model diagnostic

```{r, fig.height=9, fig.width=9, message=FALSE, warning=FALSE}
diagnostic <- augment(m_full_num_2)

a <- ggplot(diagnostic, aes(x = .fitted, y = .wtres)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE) +
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")

b <- ggplot(diagnostic, aes(x = .fitted, y = sqrt(abs(.wtres)))) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")

y <- quantile(diagnostic$.resid, c(0.25, 0.75))
x <- qnorm(c(0.25, 0.75))
slope <- diff(y) / diff(x)
int   <- y[1] - slope * x[1]     
c <- ggplot(fortify(m_full_num_2), aes(sample = .resid)) +
  geom_qq() +
  geom_abline(intercept = int, slope = slope) +
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")

cowplot::plot_grid(a, b, c)
```

Not perfect, but that'll do. 

---

### Model evaluation

```{r, warning=FALSE, message=FALSE, fig.height=6, fig.width=6}
sjp.lmer(m_full_num_2, type = "fe", prnt.plot = FALSE)$plot +
  theme_minimal(base_size = 18) +
  theme(plot.title = element_text(hjust = 0.5))
```

`r sjt.lmer(m_full_num_2, no.output = TRUE, CSS = list(css.table = "width: 100%;"))$knitr`

[Back to top]("#")

---

## Proportion of blades cleared

### Graph

```{r, warning=FALSE, message=FALSE, fig.height=6, fig.width=6}
ggplot(prop2, aes(x = Length, y = Propcleared, color = Hardness, group = Hardness)) +
  geom_jitter(alpha = 0.25, width = 0.05) +
  stat_summary(fun.y = mean, geom = "point", size = 4) +
  stat_summary(fun.y = mean, geom = "line") +
  ylab("Proportion of blades cleared") +
  scale_color_brewer("Hardness", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```

---

### Model selection 

Note: a first pass at the model showed that the residuals weren't very clean. 
I used a square root transformation to correct some of it. 

```{r, warning=FALSE, message=FALSE}
m_full_prop_2 <- lmer(sqrt(Propcleared) ~ Length * Hardness + (1 | ColonyID), data = prop2, REML = FALSE)
m_no_interaction_prop_2 <- lmer(sqrt(Propcleared) ~ Length + Hardness + (1 | ColonyID), data = prop2, REML = FALSE)
anova(m_full_prop_2, m_no_interaction_prop_2)
```

No statistical difference between the full model and the model without interaction. 
We'll continue with the model without interaction. 

---

### Model diagnostic

```{r, fig.height=9, fig.width=9, message=FALSE, warning=FALSE}
diagnostic <- augment(m_no_interaction_prop_2)

a <- ggplot(diagnostic, aes(x = .fitted, y = .wtres)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE) +
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")

b <- ggplot(diagnostic, aes(x = .fitted, y = sqrt(abs(.wtres)))) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")

y <- quantile(diagnostic$.resid, c(0.25, 0.75))
x <- qnorm(c(0.25, 0.75))
slope <- diff(y) / diff(x)
int   <- y[1] - slope * x[1]     
c <- ggplot(fortify(m_no_interaction_prop_2), aes(sample = .resid)) +
  geom_qq() +
  geom_abline(intercept = int, slope = slope) +
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")

cowplot::plot_grid(a, b, c)
```

Not perfect, but that'll do. 

---

### Model evaluation

```{r, warning=FALSE, message=FALSE, fig.height=6, fig.width=6}
sjp.lmer(m_no_interaction_prop_2, type = "fe", prnt.plot = FALSE)$plot +
  theme_minimal(base_size = 18) +
  theme(plot.title = element_text(hjust = 0.5))
```

`r sjt.lmer(m_no_interaction_prop_2, no.output = TRUE, CSS = list(css.table = "width: 100%;"))$knitr`

[Back to top]("#")

---





# Day 4

## Number of blades cleared

### Graph

```{r, warning=FALSE, message=FALSE, fig.height=6, fig.width=6}
ggplot(num4, aes(x = Length, y = Numbercleared, color = Hardness, group = Hardness)) +
  geom_jitter(alpha = 0.25, width = 0.05) +
  stat_summary(fun.y = mean, geom = "point", size = 4) +
  stat_summary(fun.y = mean, geom = "line") +
  ylab("Number of blades cleared") +
  scale_color_brewer("Hardness", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```

---

### Model selection 

Note: a first pass at the model showed that the residuals weren't very clean. 
I used a square root transformation to correct some of it. 

```{r, warning=FALSE, message=FALSE}
m_full_num_4 <- lmer(sqrt(Numbercleared) ~ Length * Hardness + (1 | ColonyID), data = num4, REML = FALSE)
m_no_interaction_num_4 <- lmer(sqrt(Numbercleared) ~ Length + Hardness + (1 | ColonyID), data = num4, REML = FALSE)
anova(m_full_num_4, m_no_interaction_num_4)
```

No statistical difference between the full model and the model without interaction. 
We'll continue with the model without interaction.

---

### Model diagnostic

```{r, fig.height=9, fig.width=9, message=FALSE, warning=FALSE}
diagnostic <- augment(m_no_interaction_num_4)

a <- ggplot(diagnostic, aes(x = .fitted, y = .wtres)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE) +
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")

b <- ggplot(diagnostic, aes(x = .fitted, y = sqrt(abs(.wtres)))) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")

y <- quantile(diagnostic$.resid, c(0.25, 0.75))
x <- qnorm(c(0.25, 0.75))
slope <- diff(y) / diff(x)
int   <- y[1] - slope * x[1]     
c <- ggplot(fortify(m_no_interaction_num_4), aes(sample = .resid)) +
  geom_qq() +
  geom_abline(intercept = int, slope = slope) +
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")

cowplot::plot_grid(a, b, c)
```

Not perfect, but that'll do. 

---

### Model evaluation

```{r, warning=FALSE, message=FALSE, fig.height=6, fig.width=6}
sjp.lmer(m_no_interaction_num_4, type = "fe", prnt.plot = FALSE)$plot +
  theme_minimal(base_size = 18) +
  theme(plot.title = element_text(hjust = 0.5))
```

`r sjt.lmer(m_no_interaction_num_4, no.output = TRUE, CSS = list(css.table = "width: 100%;"))$knitr`

[Back to top]("#")

---

## Proportion of blades cleared

### Graph

```{r, warning=FALSE, message=FALSE, fig.height=6, fig.width=6}
ggplot(prop4, aes(x = Length, y = Propcleared, color = Hardness, group = Hardness)) +
  geom_jitter(alpha = 0.25, width = 0.05) +
  stat_summary(fun.y = mean, geom = "point", size = 4) +
  stat_summary(fun.y = mean, geom = "line") +
  ylab("Proportion of blades cleared") +
  scale_color_brewer("Hardness", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```

---

### Model selection 

Note: a first pass at the model showed that the residuals weren't very clean. 
I used a square root transformation to correct some of it. 

```{r, warning=FALSE, message=FALSE}
m_full_prop_4 <- lmer(sqrt(Propcleared) ~ Length * Hardness + (1 | ColonyID), data = prop4, REML = FALSE)
m_no_interaction_prop_4 <- lmer(sqrt(Propcleared) ~ Length + Hardness + (1 | ColonyID), data = prop4, REML = FALSE)
anova(m_full_prop_4, m_no_interaction_prop_4)
```

No statistical difference between the full model and the model without interaction. 
We'll continue with the model without interaction. 

---

### Model diagnostic

```{r, fig.height=9, fig.width=9, message=FALSE, warning=FALSE}
diagnostic <- augment(m_no_interaction_prop_4)

a <- ggplot(diagnostic, aes(x = .fitted, y = .wtres)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE) +
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")

b <- ggplot(diagnostic, aes(x = .fitted, y = sqrt(abs(.wtres)))) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")

y <- quantile(diagnostic$.resid, c(0.25, 0.75))
x <- qnorm(c(0.25, 0.75))
slope <- diff(y) / diff(x)
int   <- y[1] - slope * x[1]     
c <- ggplot(fortify(m_no_interaction_prop_4), aes(sample = .resid)) +
  geom_qq() +
  geom_abline(intercept = int, slope = slope) +
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")

cowplot::plot_grid(a, b, c)
```

Not perfect, but that'll do. 

---

### Model evaluation

```{r, warning=FALSE, message=FALSE, fig.height=6, fig.width=6}
sjp.lmer(m_no_interaction_prop_4, type = "fe", prnt.plot = FALSE)$plot +
  theme_minimal(base_size = 18) +
  theme(plot.title = element_text(hjust = 0.5))
```

`r sjt.lmer(m_no_interaction_prop_4, no.output = TRUE, CSS = list(css.table = "width: 100%;"))$knitr`

[Back to top]("#")

---


