---
title: "Notebook #1"
author: "Simon Garnier"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    css: custom.css
    fig_caption: true
    fig_width: 9
    fig_height: 6
    number_sections: true
    theme: cosmo
    toc: true
---

--- 

# Set analysis environment up

```{r, 'setup', message=FALSE, cache=TRUE}
# load required libraries
if (!require(pacman)) install.packages("pacman")
p_load(readr, dplyr, ggplot2, trailClearing)

# load data
dat <- read_csv("../../data/data.csv") %>%
  mutate(Treatment = ifelse(Treatment == "LS", "15P", 
                            ifelse(Treatment == "LH", "15C", 
                                   ifelse(Treatment == "SS", "5P", "5C"))))

dat_15 <- filter(dat, Treatment == "15P" | Treatment == "15C")
dat_5 <- filter(dat, Treatment == "5P" | Treatment == "5C")

# calculate percolation distribution for full grids
full_grid_15 <- makeFullGrid(nrow = 20, ncol = 15)
distr_full_15 <- percolate2(full_grid_15, n = 100000)

full_grid_5 <- makeFullGrid(nrow = 20, ncol = 5)
distr_full_5 <- percolate2(full_grid_5, n = 100000)
```

[Back to top]("#")

---

# Percolation distributions for each sample

```{r, 'perc_distr', message=FALSE, cache=TRUE}
grids_15 <- group_by(dat_15, Treatment, Rep, Day) %>%
  do(grid = makeGrid(., nrow = 20, ncol = 15))

grids_5 <- group_by(dat_5, Treatment, Rep, Day) %>%
  do(grid = makeGrid(., nrow = 20, ncol = 5))

perc_distr_15 <- ungroup(grids_15) %>% group_by(Treatment, Rep, Day) %>%
  do(perc = percolate2(.$grid[[1]], n = 100000))

perc_distr_5 <- ungroup(grids_5) %>% group_by(Treatment, Rep, Day) %>%
  do(perc = percolate2(.$grid[[1]], n = 100000))

th_min_15 <- 30
th_max_15 <- max(c(max(distr_full_15), sapply(perc_distr_15$perc, function(x) max(x)))) - th_min_15
th_min_5 <- 10
th_max_5 <- max(c(max(distr_full_5), sapply(perc_distr_5$perc, function(x) max(x)))) - th_min_5

perc_distr_stats_15 <- ungroup(perc_distr_15) %>% group_by(Treatment, Rep, Day) %>%
  mutate(median = median(perc[[1]]), min = min(perc[[1]]), max = max(perc[[1]])) %>%
  mutate(median_norm = (median - th_min_15) / th_max_15,
         min_norm = (min - th_min_15) / th_max_15,
         max_norm = (max - th_min_15) / th_max_15)

perc_distr_stats_5 <- ungroup(perc_distr_5) %>% group_by(Treatment, Rep, Day) %>%
  mutate(median = median(perc[[1]]), min = min(perc[[1]]), max = max(perc[[1]])) %>%
  mutate(median_norm = (median - th_min_5) / th_max_5,
         min_norm = (min - th_min_5) / th_max_5,
         max_norm = (max - th_min_5) / th_max_5)

perc_distr_stats <- rbind_list(perc_distr_stats_15, perc_distr_stats_5) %>%
  mutate(Treatment = factor(Treatment, levels = c("5P", "15P", "5C", "15C")))
```

```{r, 'perc_distr_1', message=FALSE, cache=TRUE}
ggplot(perc_distr_stats, aes(x = Treatment, y = median, fill = factor(Day), color = factor(Day))) +
  geom_boxplot(alpha = 0.5) +
  ylab("Median percolation time") + 
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```
<p class="caption">**Figure 1:** Median percolation time (in simulation steps) 
as a function of treatment (5P, 15P, 5C, and 15C) and observation day (green: 
day 2; orange: day 4). For each experimental replicate, 100,000 percolation
simulations were performed and the median percolation time for each replicate is
reported here. </p>


```{r, 'perc_distr_2', message=FALSE, cache=TRUE}
ggplot(perc_distr_stats, aes(x = Treatment, y = median_norm, fill = factor(Day), color = factor(Day))) +
  geom_boxplot(alpha = 0.5) +
  ylim(0, NA) + 
  ylab("Normalized median percolation time") + 
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```
<p class="caption">**Figure 2:** Normalized median percolation time (in 
simulation steps) as a function of treatment (5P, 15P, 5C, and 15C) and 
observation day (green: day 2; orange: day 4). For each experimental replicate, 
100,000 percolation simulations were performed. The resulting percolation times 
were then normalized between the theoretical minimum percolation time (10 steps
for the 5P and 5C setups; 30 steps for the 15P and 15C setups) and the maximum 
observed time in the full grid simulations. The normalized median percolation 
time for each replicate is reported here. </p>


```{r, 'perc_distr_3', message=FALSE, cache=TRUE}
ggplot(perc_distr_stats, aes(x = Treatment, y = min, fill = factor(Day), color = factor(Day))) +
  geom_boxplot(alpha = 0.5) +
  ylab("Minimum percolation time") + 
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```
<p class="caption">**Figure 3:** Minimum percolation time (in simulation steps) 
as a function of treatment (5P, 15P, 5C, and 15C) and observation day (green: 
day 2; orange: day 4). For each experimental replicate, 100,000 percolation
simulations were performed and the minimum percolation time for each replicate 
is reported here. </p>


```{r, 'perc_distr_4', message=FALSE, cache=TRUE}
ggplot(perc_distr_stats, aes(x = Treatment, y = min_norm, fill = factor(Day), color = factor(Day))) +
  geom_boxplot(alpha = 0.5) +
  ylim(0, NA) + 
  ylab("Normalized minimum percolation time") + 
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```
<p class="caption">**Figure 4:** Normalized minimum percolation time (in 
simulation steps) as a function of treatment (5P, 15P, 5C, and 15C) and 
observation day (green: day 2; orange: day 4). For each experimental replicate, 
100,000 percolation simulations were performed. The resulting percolation times 
were then normalized between the theoretical minimum percolation time (10 steps
for the 5P and 5C setups; 30 steps for the 15P and 15C setups) and the maximum 
observed time in the full grid simulations. The normalized minimum percolation 
time for each replicate is reported here. </p>


```{r, 'perc_distr_5', message=FALSE, cache=TRUE}
ggplot(perc_distr_stats, aes(x = Treatment, y = max, fill = factor(Day), color = factor(Day))) +
  geom_boxplot(alpha = 0.5) +
  ylab("Maximum percolation time") + 
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```
<p class="caption">**Figure 5:** Maximum percolation time (in simulation steps) 
as a function of treatment (5P, 15P, 5C, and 15C) and observation day (green: 
day 2; orange: day 4). For each experimental replicate, 100,000 percolation
simulations were performed and the maximum percolation time for each replicate 
is reported here. </p>

```{r, 'perc_distr_6', message=FALSE, cache=TRUE}
ggplot(perc_distr_stats, aes(x = Treatment, y = max_norm, fill = factor(Day), color = factor(Day))) +
  geom_boxplot(alpha = 0.5) +
  ylim(0, NA) + 
  ylab("Normalized maximum percolation time") + 
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```
<p class="caption">**Figure 6:** Normalized maximum percolation time (in 
simulation steps) as a function of treatment (5P, 15P, 5C, and 15C) and 
observation day (green: day 2; orange: day 4). For each experimental replicate, 
100,000 percolation simulations were performed. The resulting percolation times 
were then normalized between the theoretical minimum percolation time (10 steps
for the 5P and 5C setups; 30 steps for the 15P and 15C setups) and the maximum 
observed time in the full grid simulations. The normalized maximum percolation 
time for each replicate is reported here. </p>


[Back to top]("#")

---

# Overlap between percolation distributions of full and experimental grids

```{r, 'overlap', message=FALSE, cache=TRUE}
ovrlp_15 <- ungroup(perc_distr_15) %>% group_by(Treatment, Rep, Day) %>%
  mutate(overlap = overlap(distr_full_15, perc[[1]]))

ovrlp_5 <- ungroup(perc_distr_5) %>% group_by(Treatment, Rep, Day) %>%
  mutate(overlap = overlap(distr_full_5, perc[[1]]))

ovrlp <- rbind_list(ovrlp_15, ovrlp_5) %>%
  mutate(Treatment = factor(Treatment, levels = c("5P", "15P", "5C", "15C")))

ggplot(ovrlp, aes(x = Treatment, y = overlap * 100, fill = factor(Day), color = factor(Day))) +
  geom_boxplot(alpha = 0.5) +
  ylab("% overlap with full grid") + 
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```
<p class="caption">**Figure 7:** Percentage overlap between the distributions of 
the simulated percolation times for the observed grids and the full grid. For 
each experimental replicate, 100,000 percolation simulations were performed and
the resulting distribution was compared to the distribution of 100,000 simulated
percolation time for a full grid (i.e. uncut). The percentage overlap between
these two distributions is reported here. An overlap of 100% means that the 2 
distributions are exactly identical. An overlap of 0% means that the 2 
distributions are completely separated from each other. </p>


[Back to top]("#")

---

# Comparison with random cutting

```{r, 'random', message=FALSE, cache=TRUE}
rnd_dat <- read_csv("../../data/par2.csv", col_types = "nnnnn")

res <- c()
for (i in 1:nrow(ovrlp)) {
  tmp <- filter(rnd_dat, idx == i)
  res <- c(res, sum(ovrlp[i, ]$overlap > tmp$overlap) / nrow(tmp))
}

ovrlp$rnd <- res

ggplot(ovrlp, aes(x = Treatment, y = rnd * 100, fill = factor(Day), color = factor(Day))) +
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 1, position = "dodge") +
  ylab("> x % of random overlap") + 
  geom_hline(yintercept = c(5, 25, 50, 75, 95), linetype = 2, alpha = c(1, 0.25, 1, 0.25, 1)) +
  geom_vline(xintercept = c(0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4, 4.5), 
             alpha = c(0.5, 0.1, 0.5, 0.1, 0.5, 0.1, 0.5, 0.1, 0.5)) +
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```
<p class="caption">**Figure 8:**  </p>


[Back to top]("#")

---



