---
title: "Notebook #2"
author: "Simon Garnier"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    css: custom.css
    fig_caption: true
    fig_width: 9
    fig_height: 6
    number_sections: true
    theme: cosmo
    toc: true
---

--- 

# Set analysis environment up

```{r, 'setup', message=FALSE, cache=TRUE}
# load required libraries
if (!require(pacman)) install.packages("pacman")
p_load(readr, dplyr, ggplot2, trailClearing)

# load data
dat <- read_csv("../../data/data.csv") %>%
  mutate(Treatment = ifelse(Treatment == "LS", "15P", 
                            ifelse(Treatment == "LH", "15C", 
                                   ifelse(Treatment == "SS", "5P", "5C"))))

dat_15 <- filter(dat, Treatment == "15P" | Treatment == "15C")
dat_5 <- filter(dat, Treatment == "5P" | Treatment == "5C")

# calculate percolation distribution for full grids
full_grid_15 <- makeFullGrid(nrow = 20, ncol = 15)
distr_full_15 <- diag(bfSearch(full_grid_15, start = cbind(x = 1, y = 1:nrow(full_grid_15)),
                          end = cbind(x = ncol(full_grid_15), y = 1:nrow(full_grid_15))))

full_grid_5 <- makeFullGrid(nrow = 20, ncol = 5)
distr_full_5 <- diag(bfSearch(full_grid_5, start = cbind(x = 1, y = 1:nrow(full_grid_5)),
                         end = cbind(x = ncol(full_grid_5), y = 1:nrow(full_grid_5))))
```

[Back to top]("#")

---

# Percolation distributions for each sample

```{r, 'perc_distr', message=FALSE, cache=TRUE}
grids_15 <- group_by(dat_15, Treatment, Rep, Day) %>%
  do(grid = makeGrid(., nrow = 20, ncol = 15))

grids_5 <- group_by(dat_5, Treatment, Rep, Day) %>%
  do(grid = makeGrid(., nrow = 20, ncol = 5))

perc_distr_15 <- ungroup(grids_15) %>% group_by(Treatment, Rep, Day) %>%
  do(perc = diag(bfSearch(.$grid[[1]], start = cbind(x = 1, y = 1:nrow(.$grid[[1]])),
                         end = cbind(x = ncol(.$grid[[1]]), y = 1:nrow(.$grid[[1]])))))

perc_distr_5 <- ungroup(grids_5) %>% group_by(Treatment, Rep, Day) %>%
  do(perc = diag(bfSearch(.$grid[[1]], start = cbind(x = 1, y = 1:nrow(.$grid[[1]])),
                         end = cbind(x = ncol(.$grid[[1]]), y = 1:nrow(.$grid[[1]])))))

th_min_15 <- 30
th_max_15 <- max(c(max(distr_full_15), sapply(perc_distr_15$perc, function(x) max(x)))) - th_min_15
th_min_5 <- 10
th_max_5 <- max(c(max(distr_full_5), sapply(perc_distr_5$perc, function(x) max(x)))) - th_min_5

perc_distr_stats_15 <- ungroup(perc_distr_15) %>% group_by(Treatment, Rep, Day) %>%
  mutate(median = median(perc[[1]]), mean =  mean(perc[[1]]), 
         min = min(perc[[1]]), max = max(perc[[1]])) %>%
  mutate(median_norm = (median - th_min_15) / th_max_15,
         mean_norm = (mean - th_min_15) / th_max_15,
         min_norm = (min - th_min_15) / th_max_15,
         max_norm = (max - th_min_15) / th_max_15)

perc_distr_stats_5 <- ungroup(perc_distr_5) %>% group_by(Treatment, Rep, Day) %>%
  mutate(median = median(perc[[1]]), mean = mean(perc[[1]]),
         min = min(perc[[1]]), max = max(perc[[1]])) %>%
  mutate(median_norm = (median - th_min_5) / th_max_5,
         mean_norm = (mean - th_min_5) / th_max_5,
         min_norm = (min - th_min_5) / th_max_5,
         max_norm = (max - th_min_5) / th_max_5)

perc_distr_stats <- rbind_list(perc_distr_stats_15, perc_distr_stats_5) %>%
  mutate(Treatment = factor(Treatment, levels = c("5P", "15P", "5C", "15C")))
```

```{r, 'perc_distr_1a', message=FALSE, cache=TRUE}
ggplot(perc_distr_stats, aes(x = Treatment, y = median, fill = factor(Day), color = factor(Day))) +
  geom_boxplot(alpha = 0.5) +
  ylab("Median percolation time") + 
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```
<p class="caption">**Figure 1a:** Median percolation time (in number of 
traversed grid cells) as a function of treatment (5P, 15P, 5C, and 15C) and 
observation day (green: day 2; orange: day 4). For each experimental replicate, 
the distribution of percolation times was obtained using an unweighted 
breadth-first search algorithm to find the shortest path between each location 
on the starting row and the closest location on the final row. The median 
percolation time for each replicate is reported here.</p>


```{r, 'perc_distr_1b', message=FALSE, cache=TRUE}
ggplot(perc_distr_stats, aes(x = Treatment, y = median_norm, fill = factor(Day), color = factor(Day))) +
  geom_boxplot(alpha = 0.5) +
  ylim(0, 1) + 
  ylab("Normalised median percolation time") + 
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```
<p class="caption">**Figure 1b:** Normalised median percolation time (in number 
of traversed grid cells) as a function of treatment (5P, 15P, 5C, and 15C) and 
observation day (green: day 2; orange: day 4). For each experimental replicate, 
the distribution of percolation times was obtained using an unweighted 
breadth-first search algorithm to find the shortest path between each location 
on the starting row and the closest location on the final row. The resulting 
percolation times were then normalised between the theoretical minimum 
percolation time (10 steps for the 5P and 5C setups; 30 steps for the 15P and 
15C setups) and the maximum observed time in the full grid simulations. The 
normalised median percolation time for each replicate is reported here. </p>


```{r, 'perc_distr_2a', message=FALSE, cache=TRUE}
ggplot(perc_distr_stats, aes(x = Treatment, y = mean, fill = factor(Day), color = factor(Day))) +
  geom_boxplot(alpha = 0.5) +
  ylab("Mean percolation time") + 
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```
<p class="caption">**Figure 2a:** Mean percolation time (in number of 
traversed grid cells) as a function of treatment (5P, 15P, 5C, and 15C) and 
observation day (green: day 2; orange: day 4). For each experimental replicate, 
the distribution of percolation times was obtained using an unweighted 
breadth-first search algorithm to find the shortest path between each location 
on the starting row and the closest location on the final row. The mean 
percolation time for each replicate is reported here.</p>


```{r, 'perc_distr_2b', message=FALSE, cache=TRUE}
ggplot(perc_distr_stats, aes(x = Treatment, y = mean_norm, fill = factor(Day), color = factor(Day))) +
  geom_boxplot(alpha = 0.5) +
  ylim(0, 1) + 
  ylab("Normalised mean percolation time") + 
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")

perc_distr_stats %>%
  group_by(Treatment, Day) %>%
  summarize(mean = mean(mean_norm),
            sd = sd(mean_norm),
            median = median(mean_norm),
            min = min(mean_norm),
            max = max(mean_norm),
            iqr = IQR(mean_norm))
```
<p class="caption">**Figure 2b:** Normalised mean percolation time (in number 
of traversed grid cells) as a function of treatment (5P, 15P, 5C, and 15C) and 
observation day (green: day 2; orange: day 4). For each experimental replicate, 
the distribution of percolation times was obtained using an unweighted 
breadth-first search algorithm to find the shortest path between each location 
on the starting row and the closest location on the final row. The resulting 
percolation times were then normalised between the theoretical minimum 
percolation time (10 steps for the 5P and 5C setups; 30 steps for the 15P and 
15C setups) and the maximum observed time in the full grid simulations. The 
normalised mean percolation time for each replicate is reported here. </p>


```{r, 'perc_distr_3a', message=FALSE, cache=TRUE}
ggplot(perc_distr_stats, aes(x = Treatment, y = min, fill = factor(Day), color = factor(Day))) +
  geom_boxplot(alpha = 0.5) +
  ylab("Minimum percolation time") + 
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```
<p class="caption">**Figure 3a:** Minimum percolation time (in number of 
traversed grid cells) as a function of treatment (5P, 15P, 5C, and 15C) and 
observation day (green: day 2; orange: day 4). For each experimental replicate, 
the distribution of percolation times was obtained using an unweighted 
breadth-first search algorithm to find the shortest path between each location 
on the starting row and the closest location on the final row. The minimum 
percolation time for each replicate is reported here.</p>


```{r, 'perc_distr_3b', message=FALSE, cache=TRUE}
ggplot(perc_distr_stats, aes(x = Treatment, y = min_norm, fill = factor(Day), color = factor(Day))) +
  geom_boxplot(alpha = 0.5) +
  ylim(0, 1) + 
  ylab("Normalised minimum percolation time") + 
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```
<p class="caption">**Figure 3b:** Normalised minimum percolation time (in number 
of traversed grid cells) as a function of treatment (5P, 15P, 5C, and 15C) and 
observation day (green: day 2; orange: day 4). For each experimental replicate, 
the distribution of percolation times was obtained using an unweighted 
breadth-first search algorithm to find the shortest path between each location 
on the starting row and the closest location on the final row. The resulting 
percolation times were then normalised between the theoretical minimum 
percolation time (10 steps for the 5P and 5C setups; 30 steps for the 15P and 
15C setups) and the maximum observed time in the full grid simulations. The 
normalised minimum percolation time for each replicate is reported here. </p>


```{r, 'perc_distr_4a', message=FALSE, cache=TRUE}
ggplot(perc_distr_stats, aes(x = Treatment, y = max, fill = factor(Day), color = factor(Day))) +
  geom_boxplot(alpha = 0.5) +
  ylab("Maximum percolation time") + 
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```
<p class="caption">**Figure 4a:** Maximum percolation time (in number of 
traversed grid cells) as a function of treatment (5P, 15P, 5C, and 15C) and 
observation day (green: day 2; orange: day 4). For each experimental replicate, 
the distribution of percolation times was obtained using an unweighted 
breadth-first search algorithm to find the shortest path between each location 
on the starting row and the closest location on the final row. The maximum 
percolation time for each replicate is reported here.</p>


```{r, 'perc_distr_4b', message=FALSE, cache=TRUE}
ggplot(perc_distr_stats, aes(x = Treatment, y = max_norm, fill = factor(Day), color = factor(Day))) +
  geom_boxplot(alpha = 0.5) +
  ylim(0, 1) + 
  ylab("Normalised maximum percolation time") + 
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```
<p class="caption">**Figure 4b:** Normalised maximum percolation time (in number 
of traversed grid cells) as a function of treatment (5P, 15P, 5C, and 15C) and 
observation day (green: day 2; orange: day 4). For each experimental replicate, 
the distribution of percolation times was obtained using an unweighted 
breadth-first search algorithm to find the shortest path between each location 
on the starting row and the closest location on the final row. The resulting 
percolation times were then normalised between the theoretical minimum 
percolation time (10 steps for the 5P and 5C setups; 30 steps for the 15P and 
15C setups) and the maximum observed time in the full grid simulations. The 
normalised maximum percolation time for each replicate is reported here. </p>


[Back to top]("#")

---

# Overlap between percolation distributions of full and experimental grids

```{r, 'overlap', message=FALSE, cache=TRUE}
ovrlp_15 <- ungroup(perc_distr_15) %>% group_by(Treatment, Rep, Day) %>%
  mutate(overlap = overlap(distr_full_15, perc[[1]]))

ovrlp_5 <- ungroup(perc_distr_5) %>% group_by(Treatment, Rep, Day) %>%
  mutate(overlap = overlap(distr_full_5, perc[[1]]))

ovrlp <- rbind_list(ovrlp_15, ovrlp_5) %>%
  mutate(Treatment = factor(Treatment, levels = c("5P", "15P", "5C", "15C")))

ggplot(ovrlp, aes(x = Treatment, y = overlap * 100, fill = factor(Day), color = factor(Day))) +
  geom_boxplot(alpha = 0.5) +
  ylab("% overlap with full grid") + 
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```
<p class="caption">**Figure 5:** Percentage overlap between the distributions of 
the percolation times for the observed grids and the full grid, as a function of 
treatment (5P, 15P, 5C, and 15C) and observation day (green: day 2; orange: day 
4). The distribution of percolation times was obtained using an unweighted 
breadth-first search algorithm to find the shortest path between each location 
on the starting row and the closest location on the final row. The percentage 
overlap between these two distributions is reported here. An overlap of 100% 
means that the 2 distributions are exactly identical. An overlap of 0% means 
that the 2 distributions are completely separated from each other. </p>


[Back to top]("#")

---

# Comparison with random cutting

```{r, 'random_a', message=FALSE, cache=TRUE}
rnd_dat <- read_csv("../../data/par4.csv", col_types = "nnnnnn")

res <- c()
for (i in 1:nrow(ovrlp)) {
  tmp <- filter(rnd_dat, idx == i)
  # res <- c(res, sum(ovrlp[i, ]$overlap > tmp$overlap) / nrow(tmp))
  # res <- c(res, sum(perc_distr_stats[i, ]$mean > tmp$mean) / nrow(tmp))
  if (length(unique(tmp$mean)) > 1) {
    f <- ecdf(tmp$mean)
    res <- c(res, f(perc_distr_stats[i, ]$mean))
  } else {
    if ((perc_distr_stats[i, ]$Treatment == "5P") | (perc_distr_stats[i, ]$Treatment == "5C")) {
      if (unique(tmp$mean) == th_min_5) {
        res <- c(res, 0)
      } else {
        res <- c(res, 1)
      }
    } else {
      if (unique(tmp$mean) == th_min_15) {
        res <- c(res, 0)
      } else {
        res <- c(res, 1)
      }
    }
  }
}

rnd <- select(perc_distr_stats, Treatment, Day) %>%
  mutate(rnd = res) 

ggplot(rnd, aes(x = Treatment, y = rnd * 100, fill = factor(Day), color = factor(Day))) +
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 1, position = "dodge") +
  ylab("Percentile of random distribution") + 
  geom_hline(yintercept = c(5, 25, 50, 75, 95), linetype = 2, alpha = c(1, 0.25, 1, 0.25, 1)) +
  geom_vline(xintercept = c(0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4, 4.5), 
             alpha = c(0.5, 0.1, 0.5, 0.1, 0.5, 0.1, 0.5, 0.1, 0.5)) +
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```
<p class="caption">**Figure 6:** Mean percolation time of each replicate 
relative to the distribution of mean percolation times for 10,000 simulated 
random grids, as a function of treatment (5P, 15P, 5C, and 15C) and observation 
day (green: day 2; orange: day 4). For each replicate, we simulated 10,000 
random grids containing the same number of cut grass bits. We then calculated 
the mean percolation time for each of these random grids using the same method 
as for the experimental replicates. Finally, we computed the theoretical 
percentile of the random distribution in which each experimental mean 
percolation time would fall in. Note that in the case when no grass was cut, the 
theoretical percentile was set to 100. In the case when all grass bits were cut, 
the theoretical percentile was set to 0. </p>


```{r, 'random_b', message=FALSE, cache=TRUE}
score <- mutate(rnd, score = ifelse(res >= 0.95, 1, ifelse(res <= 0.05, -1, 0))) %>%
  group_by(Treatment, Day) %>%
  summarize(score = sum(score == -1) / length(score))

ggplot(score, aes(x = Treatment, y = score * 100, fill = factor(Day), color = factor(Day))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), width = 0.75, alpha = 0.5) +
  ylab("% better than random") + 
  scale_y_continuous(breaks = c(0, 10, 20, 30, 40, 50)) + 
  geom_vline(xintercept = c(0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4, 4.5), 
             alpha = c(0.5, 0.1, 0.5, 0.1, 0.5, 0.1, 0.5, 0.1, 0.5)) +
  scale_fill_brewer("Day", palette = "Dark2") + 
  scale_color_brewer("Day", palette = "Dark2") + 
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```
<p class="caption">**Figure 7:** Percentage of experimental replicates 
performing better than randomly cutting grass bits, as a function of treatment 
(5P, 15P, 5C, and 15C) and observation day (green: day 2; orange: day 4). For 
each condition, we counted the number of replicates which mean percolation time
was lower than 5% of the mean percolation times of the corresponding randomly 
generated grids (see Figure 6). </p>


[Back to top]("#")

---
